Species motif participation provides unique information about species risk of extinction

### 2023-05-11 manuscript/reorder_main_text.tex
Updates from Overleaf

\author{Alyssa R. Cirtwill$^{1*\dagger}$, Anna \r{A}kesson$^{2*}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{2}$} 
\date{
\small$^1$Spatial Foodweb Ecology Group\\
Research Centre for Ecological Change\\
Organismal and Evolutionary Biology Research Programm\\
Faculty of Biological and Environmental Sciences\\
University of Helsinki, Helsinki, Finland\\
\medskip
\small$^2$Department of Theoretical Biology, Chemistry, and Physics\\ 
Link\"{o}ping University, Link\"{o}ping, Sweden\\
\medskip
\small$^3$ BioFrontiers Institute\\
University of Colorado, Boulder, Boulder, U.S.A.\\
\medskip
* Joint first authors\\
\medskip
$^\dagger$ Corresponding author email: alyssa.cirtwill@gmail.com\\
}

@@

    \indent{1. Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.} \\
    2. Given known relationships between coarse measures of a consumer's position in a food web and its ability to persist following disturbance, we expect that a species' participation in three-species motifs --meso-scale structures which provide information on a species' direct and indirect interactions-- will also be related to its probability of persistence.\\
    3. We show that a consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.
    4. However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.
    5.

@@

    The loss of a species in a food web can, due to the mutual dependencies among species,  trigger a cascade of additional (secondary) extinctions. 
    A species' risk of going secondarily extinct is known to depend on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
    For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{binzer2011susceptibility, Eklof2006}.
    Species with low in-degree are more vulnerable to extinction because they have few alternative resources if a prey species is lost, while those with high trophic levels are vulnerable to the loss of any species along the path from basal resource(s) to the focal species.
    As well as these simple measures of food-web position, it has also been proposed that a species' participation in \emph{meso-scale structures} (which take into account the species' direct and indirect interactions) are related to its persistence.

    One way to describe a species' participation in such meso-scale structures is using its participation in the set of \emph{motifs} of a given size~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
    Motifs --unique patterns of $n$ interacting species-- correspond to different sets of direct and indirect interactions and provide a holistic description of a species' place in a network~\citep{Stouffer2007,Stouffer2012}.
    Like degree and trophic level, a species' \emph{motif participation} (the vector of frequencies with which it appears in each motif) can easily be calculated once the structure of the network is known.
    However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the positions of focal species' interaction partners.
    Two species might both have two prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
    Importantly, since specialists are generally more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
    Therefore, a species' motif participation profile captures the indirect interactions that affect the species, thereby capturing important structures that are missed by larger- or smaller-scale descriptions of the network.
    As such, motif participation may complement degree and trophic level to more fully predict species' extinction risk.

    Here, we focus on persistence after the loss of plants or other basal resources, i.e., the effects of bottom-up disturbances.
    The plant community is the foundation upon which the myriad species in food webs depend for their survival, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
    Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
    It is therefore critically important to understand how bottom-up effects of disturbance affect the rest of the food web.
    As well as identifying identifying patterns of motif participation that are associated with greater or lesser ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
    For example, participation in one motif might be associated with greater probability of extinction after a mild disturbance but be associated with less risk of extinction when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances but also provide the species with alternatives when disturbances are strong.

    %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
    We define motif participation using three-species motifs, which are the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
    Four of these motifs, namely the three-species chain, apparent competition, direct competition, and omnivory, are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, Borrelli2015a, giling2019plant}.
    These four motifs also have clear ecological interpretations.
    The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
    The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers (e.g., by depletion of the shared resource). 
    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey (e.g., by boosting predator populations) in a top-down context, but in a bottom-up context only includes direct interactions.
    Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator.

@@

        We also calculated two simpler measures of a species' role within its community.
        In-degree is a species' number of prey.
        Trophic level (shortest trophic level; STL) is the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}).
        Both simple role measures were calculated in R~\citep{R} using the same algorithm as~\citet{Eklof2013}.

        Finally, to provide context for species' motif participation, we also calculated network ``motif profiles'': four-dimensional vectors of the number of each three-species motif in each network~\citep{Stouffer2012}.
        Together with network size and connectance, these motif profiles describe the structure of the whole network rather than how a particular species participates in the community.
        To separate differences in motif structure from differences in network size or connectance (larger and more-connected networks contain more motifs), we normalized motif profiles by dividing the count of each motif by the total count of all motifs. 
        We calculated species' motif participation using the Python package \emph{pymfinder}~\citep{pymfinder}.

@@

        Bayesian networks also require the creation of acyclic networks so that a strict bottom-up sequence can be followed. 
        This means that cycles where species A eats species B and vice versa or where A eats B eats C eats A must be broken (see \emph{Appendix SX} for details).
        Since cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited~\citep{Allesina2009functional}, this change will not strongly affect consumers' extinction risk in our simulations.

        The removal of cycles also affects which motifs can appear in the network. 
        Any motif which contains a cycle is impossible in an acyclic network.
        This means that only four motifs can appear: apparent competition, direct competition, omnivory, and the three-species chain).
        Fortunately, they are also the motifs that are most common in empirical networks (\citealp[]{Stouffer2007}, have been most studied in isolation (e.g.,~\citep{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.

@@

        We first tested 1) how consumer species persistence varied with motif participation and the severity of disturbance.
        For context, we also test 2) how a consumer's persistence varies with other measures of network structure and a species' place within it.
        As different measures of network structure are not independent since all describe the same network, we also 3) relate species' motif participation to large-scale network structure and to species in-degree and trophic level.
        All analyses are briefly described below and were performed using  R~\citep{R} package.% \emph{lmerTest}~\citep{lmerTest}.
        % , \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.
        See Appendix \emph{S3} for details.

@@

    Identifying the species that are most and least likely to persist after a disturbance is a high priority for conservation efforts~\citep{Bottrilletal2008}. 
    Network structural properties such as connectance and measures of a species' position in the network such as degree and trophic level are known to affect species' vulnerability~\citep{Dunne2002, Eklof2006, PascualDunne2006,prill2005dynamic, bascompte2005simple,curtsdotter2011robustness, dunne2009cascading}, but are relatively coarse metrics.
    Defining a species' network role based on its participation in meso-scale motifs provides more nuanced information and may also predict extinction risk~\citep{Stouffer2007,Stouffer2010b,Simmons2019}.
    Here we tested how the persistence of a consumer in a food web depends on its participation in four different motifs with different levels of disturbance to basal resources.  
    We found that motif participation was related to a consumer's probability of persistence in ways that can't be fully explained by simpler measures of network structure. 
    Species participating more frequently in the direct competition motif and less frequently in the omnivory motif are more likely to persist, while the effects of participating in the apparent competition or three-species chain motifs depend on the strength of disturbance.

    Disturbance strength is often neglected when predicting risk of extinction, but proved to be extremely important in our results.
    Several of the relationships we demonstrated between probability of persistence and species roles changed direction when moving from weak to strong disturbance.
    These simulated results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
    For example, at low levels of disturbance a high degree may be protective since species can compensate if one or a few prey are lost. 
    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and have lower persistence.
    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{binzer2011susceptibility}.
    These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.

    % Apparent competition as a case study that motifs are useful
    The relationship between probability of persistence and participation in the apparent competition motif provides a good example of how motifs provide complementary information to that given by simpler measures of a species' network role.
    Species with higher trophic levels tended to participate more frequently in the apparent competition motif.
    As species with higher trophic levels generally have increased risk of going extinct~\citep{Eklof2006,Eklof2013}, we would therefore expect high frequencies of participation in the apparent competition motif to also be associated with lower probability of persistence.
    However, this is only true for low disturbance levels. 
    For high disturbance, we see the opposite relationship – higher probability of persistence with more-frequent participation in apparent competition.
    This change in the persistence-apparent competition relationship with changing disturbance is what we might expect based on the negative relationship between frequency of participating in apparent competition and in-degree (which showed the opposite trends with respect to persistence).
    However, it is not obvious when trophic level should be the most important predictor of extinction risk and when in-degree should be most important.
    Simpler measures of species' positions in a network could therefore lead us astray either because they provide conflicting predictions or because our expectations are only reasonable for specific types of disturbance.

    % Try to condense the next few bits into one section on omnivory
    Frequency of participation in the omnivory motif likewise provides information about extinction risk that is not easily obtained from in-degree and trophic level.
    Higher frequencies of participation in the omnivory motif were associated with lower probabilities of persistence, yet the frequency of the omnivory motif increased with in-degree and decreased with trophic level.
    Both high in-degree and low trophic level are associated with higher probability of persistence, except at high levels of disturbance, which would naively suggest that higher proportions of omnivory should also be associated with greater persistence. 
    Considering a species' in-degree, trophic level, \emph{and} its motif role may therefore provide extra information about its extinction risk.

    While greater frequency of participation in the omnivory motif was generally associated with reduced probability of persistence across all levels of disturbance, this trend varied strongly with overall network structure (especially connectance).
    Under weak disturbance scenarios, the effect of increased participation in the omnivory motif on consumer persistence within a single network was likely to be neutral or positive (Fig. 3). 
    At high levels of disturbance, however, participating in the omnivory motif was more likely to reduce the probability of persistence as in~\citet{McLeod2021}.
    This is consistent with other studies which have found the effect of omnivory to be highly context-dependent~\citep{bascompte2005simple, Monteiro2016}.

@@

    % Caveats - expand on top-down effects a bit to make R2 happy
    There are limitations to the approach presented here.
    Most importantly, the Bayesian network framework operates on a strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
    Here our focus was to analyze bottom-up affects following disturbances to the basal resources, in which case the Bayesian network approach have been shown to capture a majority of the secondary extinctions obtained in a fully-fledged dynamical model but doing so with much greater efficiency~\citep{Eklof2013}.
    If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
    For example, although a recent study using a dynamical model to relate motif participation to persistence after a disturbance found similar trends for omnivory, participation in the direct competition motif was not universally positive~\citep{Cirtwill2021_inprep}.
    In particular, species which frequently appeared in the prey position of the direct competition motif tended to have shorter persistence times after disturbance in that study - possibly due to top-down effects from the pair of predators involved.
    However, as these two studies applied different types of disturbances, further study would be required to disentangle the effects of top-down and bottom-up effects on species with different motif participation, and how each relates to the type and strength of disturbance.
    For this preliminary study, we have also neglected species traits.
    As traits such as body size are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000} in empirical webs, this is also a potentially useful addition to future work.

>>>>

\author{Alyssa R. Cirtwill$^{1*\dagger}$, Anna \r{A}kesson$^{2*}$, Kate L. Wootton$^{3}$, Anna Ekl\"{o}f$^{2}$} 
\date{
\small$^1$Spatial Foodweb Ecology Group\\
Research Centre for Ecological Change\\
Organismal and Evolutionary Biology Research Programm\\
Faculty of Biological and Environmental Sciences\\
University of Helsinki, Helsinki, Finland\\
\medskip
\small$^2$Department of Theoretical Biology, Chemistry, and Physics\\ 
Link\"{o}ping University, Link\"{o}ping, Sweden\\
\medskip
\small$^3$ School of Biological Sciences\\
University of Canterbury, Christchurch, New Zealand\\
\medskip
* Joint first authors\\
\medskip
$^\dagger$ Corresponding author email: alyssa.cirtwill@gmail.com\\
}

@@

    \indent{1. Bottom-up effects of disturbances to basal resources can strongly affect the persistence of consumers, but it is difficult to predict which species will be most affected by increasing disturbance.} \\
    2. Given known relationships between coarse measures of a consumer's position in a food web and its ability to persist following disturbance, we expect that a species' participation in three-species motifs --meso-scale structures which provide information on a species' direct and indirect interactions-- will also be related to its probability of persistence.\\
    3. We show that consumer species which participate more frequently in the direct competition motif and less frequently in the omnivory motif generally have higher probability of persistence following disturbance to basal resources.\\
    4. However, both the strength of the disturbance and the overall network structure (i.e., connectance) affect the strength and direction of relationships between motif participation and persistence.\\
    5.

@@

    The loss of a species in a food web can, due to the mutual dependencies among species,  trigger a cascade of additional (secondary) extinctions. 
    A species' risk of going secondarily extinct is known to depend on its position within a food web~\citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
    For example, species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally more likely to go extinct~\citep{Allesina2012,binzer2011susceptibility, Eklof2006}.
    Species with low in-degree are more vulnerable to extinction because they have few alternative resources if a prey species is lost, while those with high trophic levels are vulnerable to the loss of any species along the path from basal resource(s) to the focal species.
    As well as these simple measures of food-web position, it has also been proposed that a species' participation in \emph{meso-scale structures} (which take into account the species' direct and indirect interactions) are related to its persistence.

    One way to describe a species' participation in such meso-scale structures is using its participation in the set of \emph{motifs} of a given size~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}.
    Motifs --unique patterns of $n$ interacting species-- correspond to different sets of direct and indirect interactions and provide a holistic description of a species' place in a network~\citep{Stouffer2007,Stouffer2012}.
    Like degree and trophic level, a species' \emph{motif participation} (the vector of frequencies with which it appears in each motif) can easily be calculated once the structure of the network is known.
    However, unlike in-degree and trophic level, which focus only on direct prey of the focal species or on the lengths of paths between the focal species and basal resources, motif participation also incorporates information on the positions of focal species' interaction partners.
    Two species might both have two prey and a trophic level of 3, but they might have quite different motif participation if one species's prey are strict specialists and the others' prey are generalists.
    Importantly, since specialists are usually more vulnerable to extinction than generalists, the predator of specialists will be more likely to lose prey and go extinct itself than will a predator of generalists.
    Therefore, a species' motif participation profile captures the indirect interactions that affect the species, thereby capturing important structures that are missed by larger- or smaller-scale descriptions of the network.
    As such, motif participation may complement degree and trophic level to more fully predict species' extinction risk.

    Here, we focus on persistence after the loss of plants or other basal resources, i.e., the effects of bottom-up disturbances.
    The plant community is the foundation upon which the myriad species in food webs depend for their survival, and loss of basal resource diversity typically causes declines in both abundance and richness of all types of consumers~\citep{scherber2010bottom,Dobson2009food, Mduma1999food, Georgiadis2007}.
    Moreover, disturbances to the plant community have been shown to affect important ecosystem properties such as primary \citep{Hector1999} and secondary production \citep{borer2012plant}, soil respiration and carbon cycling \citep{chen2019plant}, and consumer diversity \citep{scherber2010bottom, Baiser2016,li2020bottom}.
    It is therefore critically important to understand how bottom-up effects of disturbance affect the rest of the food web.
    As well as identifying patterns of motif participation that are associated with greater or lesser ability to persist after a loss of basal resources, we are particularly interested in whether the strength of  disturbance affects relationships between motif participation and persistence.
    For example, participation in one motif might be associated with greater probability of extinction after a mild disturbance but be associated with less risk of extinction when when more basal resources are lost if the direct and indirect interactions within the motif amplify the effect of small disturbances but also provide the species with alternatives when disturbances are strong.

    %There are 13 possible three-species motifs (the most commonly-used size for food web analyses)~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}, 
    We define motif participation using three-species motifs, which are the most commonly-used size for food web analyses~\citep{Milo2004,Stouffer2007,Stouffer2012,Cirtwill2015a}. 
    Four of these motifs, namely the three-species chain, apparent competition, direct competition, and omnivory, are especially interesting as they are by far the most common in empirical networks~\citep{Stouffer2007, Borrelli2015a, giling2019plant}.
    These four motifs also have clear ecological interpretations.
    The three-species chain indicates vertical flows of energy and biomass in a network and captures the possibility of indirect interactions across multiple trophic levels (e.g., between a plant and a predator of the plant's herbivore).
    The direct competition motif, where two consumers share the same resource, captures the possibility of indirect interactions between consumers (e.g., by depletion of the shared resource). 
    The apparent competition motif, where a predator consumes two resources, can include indirect effects between prey (e.g., by boosting predator populations) in a top-down context, but in a bottom-up context, such as here, only includes direct interactions.
    Finally, the omnivory motif, where a predator consumes the same resource as its prey, includes captures the potential for a plant to have both direct and indirect effects on the predator.

@@

        We also calculated two simpler measures of a species' role within its community.
        In-degree is a species' number of prey.
        Trophic level (shortest trophic level; STL) is the length of the shortest food chain between the focal species and any basal resource~\citep{Williams2004}.
        Both simple role measures were calculated in R~\citep{R} using the same algorithm as~\citet{Eklof2013}.

        Finally, to provide context for species' motif participation, we calculated network ``motif profiles'': four-dimensional vectors of the number of each three-species motif in each network~\citep{Stouffer2012}.
        Together with network size and connectance, these motif profiles describe the structure of the whole network rather than how a particular species participates in the community.
        To separate differences in motif structure from differences in network size or connectance (larger and more-connected networks contain more motifs), we normalized motif profiles by dividing the count of each motif by the total count of all motifs. 
        We calculated species' motif participation using the Python package \emph{pymfinder}~\citep{pymfinder}.

@@

        Bayesian networks also require the creation of acyclic networks so that a strict bottom-up sequence can be followed. 
        This means that cycles where species A eats species B and vice versa or where A eats B eats C eats A must be broken (see \emph{Appendix SX} for details).
        Since cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effects on network robustness are very limited~\citep{Allesina2009functional}, this change will not strongly affect consumers' extinction risk in our simulations.

        The removal of cycles also affects which motifs can appear in the network. 
        Any motif which contains a cycle is impossible in an acyclic network.
        This means that only four motifs can appear: apparent competition, direct competition, omnivory, and the three-species chain.
        Fortunately, they are also the motifs that are most common in empirical networks (\citealp[]{Stouffer2007}, have been most studied in isolation (e.g.,~\citep{Hastings1991,Holt1987,Bascompte2005,Polis1989,Zabalo2012,Lefevre2009,Holt1997,Kondoh2008,McKinnon2013,Laws2013}), have the clearest ecological interpretations and, because of these three properties, are of greatest interest to us.

@@

        We first tested 1) how consumer species persistence varied with motif participation and the severity of disturbance.
        For context, we also test 2) how a consumer's persistence varies with other measures of network structure and a species' place within it.
        As different measures of network structure are not independent since all describe the same network, we also 3) relate species' motif participation to large-scale network structure and to species in-degree and trophic level.
        All analyses are briefly described below and were performed using  R~\citep{R} package. % \emph{lmerTest}~\citep{lmerTest}.
        % , \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.
        See Appendix \emph{S3} for details.

@@

    Identifying the species that are most and least likely to persist after a disturbance is a high priority for conservation efforts~\citep{Bottrilletal2008}. 
    Network structural properties such as connectance and measures of a species' position in the network such as degree and trophic level are known to affect species' vulnerability~\citep{Dunne2002, Eklof2006, PascualDunne2006,prill2005dynamic, bascompte2005simple,curtsdotter2011robustness, dunne2009cascading}, but are relatively coarse metrics.
    Defining a species' network role based on its participation in meso-scale motifs provides more nuanced information and may also predict extinction risk~\citep{Stouffer2007,Stouffer2010b,Simmons2019,Cirtwill2018FoodWebs}, but this potential is only beginning to be explored~\citep{Cirtwill2022Oikos}.

    Here we partially fill this knowledge gap by testing how the persistence of a consumer in a food web depends on its participation in four different motifs, and how this relationship varies with different levels of disturbance to basal resources.  
    We found that motif participation was related to a consumer's probability of persistence in ways not fully explained by simpler measures of network structure. 
    Species participating more frequently in the direct competition motif and less frequently in the omnivory motif are more likely to persist, while the effects of participating in the apparent competition or three-species chain motifs depend on the strength of disturbance.

    Disturbance strength is often neglected when predicting risk of extinction, but proved important in our results.
    Several of the relationships we demonstrated between probability of persistence and species roles changed direction when moving from weak to strong disturbance.
    These simulation results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
    For example, at low levels of disturbance a high degree may be protective since species can compensate if one or a few prey are lost. 
    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and therefore have lower persistence.
    This context-dependence contrasts with the general assumption that species with more resources will be less vulnerable~\citep{Allesina2012,binzer2011susceptibility}.
    These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.

    % Apparent competition as a case study that motifs are useful
    The relationship between probability of persistence and participation in the apparent competition motif provides a good example of how motifs provide complementary information to simpler measures of a species' network role.
    Species with higher trophic levels tend to participate more frequently in the apparent competition motif and
    species with higher trophic levels generally have increased extinction risk~\citep{Eklof2006,Eklof2013}. We would therefore expect high frequencies of participation in the apparent competition motif to also be associated with lower persistence probability.
    However, this is only true for low disturbance. 
    For high disturbance, we see the opposite relationship – higher probability of persistence with more-frequent participation in apparent competition.
    This change in the persistence-apparent competition relationship across disturbance levels may be driven by the negative relationship between apparent competition and in-degree, and in-degree's negative relationship with persistence.
    However, it is not obvious when trophic level should be the most important predictor of extinction risk and when in-degree should be most important.
    Such simpler measures of species' positions in a network could therefore lead us astray either because they provide conflicting predictions or because our expectations are only reasonable for specific types or levels of disturbance.

    % Try to condense the next few bits into one section on omnivory
    Omnivory motif participation likewise provides information about extinction risk that is not easily obtained from in-degree and trophic level.
    Omnivory motif participation decreased with trophic level (as does persistence) and increased with in-degree (as does persistence, except at high disturbance). This would naively suggest that higher omnivory participation should also be associated with greater persistence, but instead we (usually) find the opposite. The exception to this pattern is found at high connectance, where increased omnivory participation was more likely to have a positive or neutral relationship with consumer persistence, especially at low disturbance levels (Fig. 3).
    %While greater frequency of participation in the omnivory motif was generally associated with reduced probability of persistence across all levels of disturbance, this trend varied strongly with overall network structure (especially connectance).
    %Under weak disturbance scenarios, the effect of increased participation in the omnivory motif on consumer persistence within a single network was likely to be neutral or positive (Fig. 3). 
    %At high levels of disturbance, however, participating in the omnivory motif was more likely to reduce the probability of persistence as in~\citet{McLeod2021}.
    This is consistent with other studies which have found the effect of omnivory to be highly context-dependent~\citep{bascompte2005simple, Monteiro2016}.
    %Considering a species' in-degree, trophic level, \emph{and} its motif role may therefore provide extra information about its extinction risk.

@@

    % Caveats - expand on top-down effects a bit to make R2 happy
    There are limitations to the approach presented here.
    Most importantly, the Bayesian network framework operates on a strict bottom-up principle and therefore cannot include top-down effects in the extinction risk calculations~\citep{Eklof2013}. 
    Here our focus was to analyze bottom-up affects following disturbances to the basal resources, in which case the Bayesian network approach captures a majority of the secondary extinctions obtained in a fully-fledged dynamical model with much greater efficiency~\citep{Eklof2013}.
    If top-down effects are important for a particular empirical system, however, our simulations may not be accurate. 
    For example, although a recent study using a dynamical model to relate motif participation to persistence after a disturbance found similar trends for omnivory, participation in the direct competition motif was not universally positive~\citep{Cirtwill2021_inprep}.
    In particular, species which frequently appeared in the prey position of the direct competition motif tended to have shorter persistence times after disturbance in that study - possibly due to top-down effects from the pair of predators involved.
    However, as these two studies applied different types of disturbances, further study would be required to disentangle the effects of top-down and bottom-up effects on species with different motif participation, and how each relates to the type and strength of disturbance.
    For this preliminary study, we have also neglected species traits.
    As traits such as body size are related to both motif participation~\citep{cirtwill2018feeding} and extinction risk~\citep{Brose2017, curtsdotter2011robustness, Cardillo2005, Purvis2000} in empirical webs, this is also a potentially useful addition to future work.
